Real-Life Algebraic File Synchronization

### 2015-10-26 2015/paper.tex
Continued with describing the new algebra

\newcommand{\setv}{\mathcal{V}}
\newcommand{\setf}{\mathcal{V}_F}
\newcommand{\setd}{\mathcal{V}_D}
\newcommand{\setb}{\mathcal{V}_\empt}
\newcommand{\setp}{\mathcal{P}}
\newcommand{\empt}{\bot}
\newcommand{\parent}{\mathtt{parent}}
\newcommand{\FS}{\Phi} % {\mathrm{FS}}

\newcommand{\cbrk}{\mathtt{break}}
\newcommand{\cbb}{\overrightarrow{\empt\empt}}
\newcommand{\cbf}{\overrightarrow{\empt{F}}}
\newcommand{\cbd}{\overrightarrow{\empt{D}}}
\newcommand{\cfb}{\overrightarrow{F\empt}}
\newcommand{\cff}{\overrightarrow{FF}}
\newcommand{\cfd}{\overrightarrow{FD}}
\newcommand{\cdb}{\overrightarrow{D\empt}}
\newcommand{\cdf}{\overrightarrow{DF}}
\newcommand{\cdd}{\overrightarrow{DD}}
\newcommand{\cxy}{\overrightarrow{XY}}

@@

\begin{document}
\begin{abstract}
Abstract goes here
\end{abstract}

@@

Let us consider what kind of information may be encoded in the commands.
A usual set of commands, based on the most frequent tools implemented by filesystems,
might be $create(p,V)$, $edit(p,V)$ and $remove(p)$ where $p\in\setp$ and $V\in\setv$.
Clearly the commands need to contain information about the state they leave the filesystem
in at the path on which they operate, that is, they need to contain the ``end'' value, $V$.
Notice, however, that the commands above also encode some information about the filesystem
{\it before} the command is applied; namely, $create$ requires that there are no files
or directories at $p$, while $edit$ and $remove$ require the opposite.

@@

We therefore propose to have a command for each pair of types.
For want of a better system, we will simply name our commands by concatenating
two of $D$, $F$ and $\empt$ with an arrow, where the left sign notes the type of value
in the filesystem before the command is applied, and the right sign notes the type
afterwards. For example, $mkdir(p,D)$ is $\cbd(p,D)$ and $rmdir(p)$ is $\cdb(p)$.

% No "move"

@@

So that we could reason about sequences of commands that break every filesystem
in the algebra, we introduce the command $\cbrk$ that simply breaks every filesystem.
We note that a command or a sequence of commands $S$ is applied to a filesystem
by prefixing the command or sequence to it:
$S\FS$; or $\cbrk\FS=\empt$; or $\Xi = \cff(p,F)\FS$, for which $\Xi(p)=F$ holds
and $\Xi(q)=\FS(q)$ for any path $q\neq p$.

@@

>>>>

\newcommand{\setv}{\mathcal{V}}
\newcommand{\setvx}[1]{\mathcal{V}_{#1}}
\newcommand{\setf}{\setvx{F}}
\newcommand{\setd}{\setvx{D}}
\newcommand{\setb}{\setvx{\empt}}
\newcommand{\setp}{\mathcal{P}}
\newcommand{\empt}{\bot}
\newcommand{\parent}{\mathtt{parent}}
\newcommand{\FS}{\Phi} % {\mathrm{FS}}
\newcommand{\GS}{\Xi} % another filesystem

\newcommand{\cbrk}{\mathtt{break}}
\newcommand{\cbb}{\overrightarrow{\empt\empt}}
\newcommand{\cbf}{\overrightarrow{\empt{F}}}
\newcommand{\cbd}{\overrightarrow{\empt{D}}}
\newcommand{\cfb}{\overrightarrow{F\empt}}
\newcommand{\cff}{\overrightarrow{FF}}
\newcommand{\cfd}{\overrightarrow{FD}}
\newcommand{\cdb}{\overrightarrow{D\empt}}
\newcommand{\cdf}{\overrightarrow{DF}}
\newcommand{\cdd}{\overrightarrow{DD}}
\newcommand{\cxy}{\overrightarrow{XY}}
\newcommand{\czw}{\overrightarrow{ZW}}

@@

\begin{document}
\maketitle
\begin{abstract}
Abstract goes here
\end{abstract}

\section{Introduction}

% Synchronization:
% Main aim: apply updates (where possible) from other replicas
% update detection -> conflict detection -> merging updates -> ordering updates -> apply updates from other replicas
% Problems: conflict detection, ordering

@@

Let us consider what kind of information may be encoded in the commands.
A usual set of commands, based on the most frequent tools implemented by filesystems,
might be $create(p,V)$, $edit(p,V)$ and $remove(p)$ where $p\in\setp$ and $V\in\setv$ (but $V\neq\empt$).
Clearly the commands need to contain information about the state they leave the filesystem
in at the path on which they operate, that is, they need to contain the ``end'' value, $V$.
Notice, however, that the commands above also encode some information about the filesystem
{\it before} the command is applied; namely, $create$ requires that there are no files
or directories at $p$, while $edit$ and $remove$ require the opposite.

@@

We therefore propose to have a command for each pair of types.
For want of a better system, we will simply name our commands by concatenating
two of $D$, $F$ and $\empt$ with an arrow, where the left sign notes the type of value
in the filesystem before the command is applied, and the right sign notes the type
afterwards. For example $mkdir(p,D)$ is $\cbd(p,D)$ and $rmdir(p)$ is $\cdb(p)$.

% TODO No "move"

@@

So that we could reason about sequences of commands that break every filesystem
in the algebra, we introduce the command $\cbrk$ that simply breaks every filesystem.
We note that a command or a sequence of commands $S$ is applied to a filesystem
by prefixing the command or sequence to it:
\begin{itemize}
\item $S\FS$
\item $\cbrk\FS=\empt$
\item $\cff(p,F)\FS = \GS$, for which $\GS(p)=F$ holds
and $\GS(q)=\FS(q)$ for any path $q\neq p$.
\end{itemize}

@@

We aim to use our algebra to define and implement algorithms for conflict detection
and ordering updates (commands) before we can apply them to all replicas.
Conflicts and ordering all happen between commands, therefore in our algebra
we will reason about sequences of commands, 
and do so based on a set of axioms about pairs of neighbouring commands.
We are interested to see what pairs of commands can be simplified or reversed
in order to see under what circumstances commands conflict or can be reordered.

Pairs of commands in general will have the form
\[ \cxy(p,V_Y) \czw(q,V_W) \]
where $X,Y,Z,W\in\{D,F,\empt\}$, $p,q\in\setp$, and values are of the appropriate type: $V_Y\in\setvx{Y}$ and $V_W\in\setvx{W}$
(although commands where $Y=\empt$ or $W=\empt$ do not require a value).

Similarly to the case of commands, in \cite{NREC:alg} we found that in order to build
a set of axioms in a sound and complete algebra we at times needed to take into account
the relationship between the paths of the commands in the pair, $p$ and $q$.
As the relationship between the paths determine how the commands affect the filesystem
and its tree property, 
and again in line with the aim of providing as much information as possible for the algebra,
we encoded this relationship in the pair of commands. The possible relationships we
need to take into account are as follows.

\begin{itemize}
\item $p=q$.
\item $\parent(p)=q$ and $\parent(q)=p$. We know that 
\item Only child
\item separate
\end{itemize}
